---
title: "Electronic Monitoring EFP Reporting"
author: "george@capecodfishermen.org"
date: "Fishing Year 2019"
output: word_document
---
  
```{r setup, include=FALSE}
library(RJSONIO)
library(XLConnect)
library(lubridate)
library(stringdist)
library(rgdal)
library(zip)
library(sp)
library(marmap)
library(devtools)
library(dplyr)
library(sf)
library(flextable)
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
options(scipen=999)
```
```{r Data Download and Processing, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## The marmap library is partly deprecated because of NOAA's website updates
## Users in the GitHub community have developed a new function, and the code
## below downloads the most recent version from Eric Pante's GitHub repo
source_url(
  url="https://raw.githubusercontent.com/ericpante/marmap/master/R/getNOAA.bathy.R"
)
## The three datasets that make up the basis of this analysis are the FY2019 VTR
## data (requested from GARFO), the FY2019 EM data (requested from Teem.Fish),
## and the FY2019 Dealer Data (requested from sector managers). The next step
## reads in each dataset individually. All data are housed in the Electronic 
## Monitoring directory of the CCCFA server. The default starting point for 
## file addresses is 
## "Electronic Monitoring/Georges Analyses/ElectronicMonitoring"
##
## VTR data is read in from a .csv file, which functions as a data frame 
VTR=read.csv(
  file="../../SMAST Science/Data/FY2019-eVTR-GARFO-20201007.csv"
)
## EM data is read in from a .json file which functions as a list of lists
EM=fromJSON(
  content="../ClosedAreaComparisons/FY19/RawData/NOAA_Submissions_2019.json"
)
## Dealer data is read in from a series of .xlsx and .xls files mailed
## from the sector managers. All of this data should be stored in the same 
## directory to enable gathering. the Sustainable Harvest Sector's manager sends
## both eVTR and dealer data in the same file, so it is important to load the 
## correct worksheet from that file
fileList=dir("../ClosedAreaComparisons/FY19/RawData/DealerData/")
Dealer=data.frame()
for(i in 1:length(fileList)){
  filename=paste0(
    "../ClosedAreaComparisons/FY19/RawData/DealerData/",
    fileList[i]
  )
  if(
    grepl(
      pattern="shs",
      x=filename
    )==FALSE
  ){
    if(
      grepl(
        pattern=".xlsx",
        x=filename
      )==TRUE
    ){
      partial=XLConnect::readWorksheetFromFile(
        file=filename,
        sheet=1
      )
    } else {
      if(
        grepl(
          pattern=".csv",
          x=filename
        )==TRUE
      ){
        partial=read.csv(
          file=filename
        )
      }
    }
  } else {
    partial=XLConnect::readWorksheetFromFile(
      file=filename,
      sheet="dealer"
    )
  }
  ## Sometimes the SIMM exports the first column with an extra space in the
  ## column name, so check for that and correct it if it exists
  colnames(partial)[1]=ifelse(
    grepl(
      pattern="Sector.Id",
      x=colnames(partial)[1]
      ),
    "Sector.Id",
    "ERROR"
  )
  Dealer=rbind(Dealer,partial)
}
## The next step is to turn the .json EM file into a dataframe to enable merging
## it with other records
## Rename the EM .json to EM_JSON and create a new empty EM data frame
EM_JSON=EM
EM=data.frame(
  VTR=as.numeric(),
  VESSEL=as.character(),
  HAUL_NO=as.numeric(),
  startTime=as.numeric(),
  endTime=as.numeric(),
  startLat=as.numeric(),
  startLon=as.numeric(),
  species=as.character(),
  count=as.numeric(),
  weight=as.numeric()
)
## Each trip is an item in a list
## For each trip
for(i in 1:length(EM_JSON)){
  ## Open the list item
  trip=EM_JSON[[i]]
  ## Each trip is a list of variables and dataframes
  ## Extract the VTR number
  vtr=trip$trip_id
  ## Extract the vessel name
  vessel=trip$vessel_name
  ## Extract the number of hauls
  hauls=trip$total_hauls
  ## Report discards haul by haul
  for(h in 1:hauls){
    haul=trip$hauls[[h]]
    startTime=haul[2]
    endTime=haul[3]
    startLat=haul[4]
    startLon=haul[5]
    ## Discards are listed by species
    discards=haul$discards
    for(d in 1:length(discards)){
      species=discards[[d]]$species
      count=discards[[d]]$count_discarded
      weight=discards[[d]]$pounds_discarded
      newline=data.frame(
        VTR=as.numeric(),
        VESSEL=as.character(),
        HAUL_NO=as.numeric(),
        startTime=as.numeric(),
        endTime=as.numeric(),
        startLat=as.numeric(),
        startLon=as.numeric(),
        species=as.character(),
        count=as.numeric(),
        weight=as.numeric()
      )
      newline[1,]=NA
      newline$VESSEL=as.character(newline$VESSEL)
      newline$species=as.character(newline$species)
      newline$VTR=vtr
      newline$VESSEL=toupper(
        as.character(vessel)
      )
      newline$HAUL_NO=h
      newline$startTime=startTime
      newline$endTime=endTime
      newline$startLat=startLat
      newline$startLon=startLon
      newline$species=toupper(
        as.character(species)
      )
      newline$count=count
      newline$weight=weight
      EM=rbind(EM,newline)
      rm(newline)
    }
  }
}
## Create interoperable data sets that have all the information of interest
## Read in the species standardization list
species=read.csv(
  "https://raw.githubusercontent.com/gamaynard/ElectronicMonitoring/master/species.csv"
)
## ---------------------------
## VTR data
iVTR=VTR[,c(
  "DATE_SAIL",
  "DATE_LAND",
  "VESSEL_PERMIT_NUM",
  "SERIAL_NUM",
  "GEARCODE",
  "GEARQTY",
  "GEARSIZE",
  "AREA",
  "LAT_DEGREE",
  "LAT_MINUTE",
  "LAT_SECOND",
  "LON_DEGREE",
  "LON_MINUTE",
  "LON_SECOND",
  "NTOWS",
  "DATETIME_HAUL_START",
  "DATETIME_HAUL_END",
  "SPECIES_ID",
  "KEPT",
  "DISCARDED",
  "PORT_LANDED"
)]
colnames(iVTR)=tolower(
  colnames(iVTR)
)
## SAILDATE should be a POSIX value
iVTR$SAILDATE=ymd_hms(
  as.character(iVTR$date_sail)
  )
## LANDDATE should be a POSIX value
iVTR$LANDDATE=ymd_hms(
  as.character(iVTR$date_land)
)
## PERMIT should be a character string
iVTR$PERMIT=as.character(
  iVTR$vessel_permit_num
)
## Check for NA values in the seconds columns of both latitude and longitude
## and if they exist, replace them with zeros
iVTR$lat_second=ifelse(
  is.na(iVTR$lat_second),
  0,
  iVTR$lat_second
)
iVTR$lon_second=ifelse(
  is.na(iVTR$lon_second),
  0,
  iVTR$lon_second
)
## All longitude values should be west of the Prime Meridian (negative)
iVTR$lon_degree=ifelse(
  iVTR$lon_degree>0,
  iVTR$lon_degree*-1,
  iVTR$lon_degree
)
## Combine degrees, minutes, and seconds into decimal degrees for both latitude
## and longitude
iVTR$LAT=iVTR$lat_degree+iVTR$lat_minute/60+iVTR$lat_second/(60^2)
iVTR$LON=iVTR$lon_degree-iVTR$lon_minute/60-iVTR$lon_second/(60^2)
## Replace Gear Codes with human-readable values
iVTR$GEAR=ifelse(iVTR$gearcode=="GNS","GILLNET",
  ifelse(iVTR$gearcode=="HND","JIG",
    ifelse(iVTR$gearcode=="LLB","LONGLINE",
      ifelse(iVTR$gearcode=="OTF","TRAWL",
        ifelse(iVTR$gearcode=="PTL","LOBSTER POT",
          iVTR$gearcode
          )
        )
      )
    )
  )
## Ensure stat areas are reported as numbers
iVTR$AREA=as.numeric(
  as.character(
    iVTR$area
  )
)
## Trim serial numbers to generate VTR numbers
iVTR$VTR=NA
iVTR$VTR=ifelse(
  nchar(iVTR$serial_num)==16,
  substr(
    iVTR$serial_num,1,14
    ),
  iVTR$VTR
)
## Any trips with paper trip reports only have an 8 character identifier and
## should be excluded from this analysis, as they are likely declared out of
## fishery
iVTR=subset(
  iVTR,
  nchar(iVTR$serial_num)==16
)
## Standardize species names
iVTR$SPECIES=NA
for(i in 1:nrow(iVTR)){
  iVTR$SPECIES[i]=as.character(
    species$AFS[
      which(
        stringsim(
          a=as.character(iVTR$species_id[i]),
          b=as.character(species$PEBKAC)
        )==max(  
          stringsim(
            a=as.character(iVTR$species_id[i]),
            b=as.character(species$PEBKAC)
          )
        )
      )[1]
      ]
  )
}
## Haul start and end times should be POSIX formatted values
iVTR$HAULSTART=ymd_hms(
  as.character(
    iVTR$datetime_haul_start
    )
  )
iVTR$HAULEND=ymd_hms(
  as.character(
    iVTR$datetime_haul_end
  )
)
## Kept and discarded weights should be numeric
iVTR$KEPT=as.numeric(
  as.character(
    iVTR$kept
    )
  )
iVTR$DISCARDED=as.numeric(
  as.character(
    iVTR$discarded
  )
)
## Some VTR records only have a "serial number" instead of a 
## ---------------------------
## EM data
## because the EM data frame is already a modification of the original data, the
## script works on it directly
## The VTR column is already a character vector (to avoid loss of leading zeros)
## The VESSEL column is already a character vector
## The HAUL_NO column is already an integer
## The startTime column needs to be converted to a POSIX value
EM$STARTTIME=ymd_hm(
  as.character(
    EM$startTime
    )
  )
## The endTime column needs to be converted to a POSIX value
EM$ENDTIME=ymd_hm(
  as.character(
    EM$endTime
  )
)
## The startLat column needs to be converted to a number
EM$STARTLAT=as.numeric(
  as.character(
    EM$startLat
  )
)
## The startLon column needs to be converted to a number
EM$STARTLON=as.numeric(
  as.character(
    EM$startLon
  )
)
## Create a standardized species column
EM$SPECIES=NA
for(i in 1:nrow(EM)){
  EM$SPECIES[i]=as.character(
    species$AFS[
      which(
        stringsim(
          a=as.character(EM$species[i]),
          b=as.character(species$PEBKAC)
        )==max(  
          stringsim(
            a=as.character(EM$species[i]),
            b=as.character(species$PEBKAC)
          )
        )
      )[1]
    ]
  )
}
## Discard Count needs to be a number
EM$DiscardCount=as.numeric(
  as.character(
    EM$count
    )
  )
## DiscardWeight needs to be a number
EM$DiscardWeight=as.numeric(
  as.character(
    EM$weight
  )
)
## ---------------------------
## Dealer data
iDealer=Dealer[,c(
  "Mri",
  "Vessel.Permit.No",
  "Vessel.Name",
  "Vessel.Reg.No",
  "Vtr.Serial.No",
  "State.Land",
  "Port.Land",
  "Species.Itis",
  "Landed.Weight",
  "Live.Weight",
  "Date.Sold"
)]
## Permit numbers should be character strings to maintain leading zeros
iDealer$PERMIT=as.character(
  iDealer$Vessel.Permit.No
  )
## Vessel names should be all caps
iDealer$VESSEL=toupper(
  as.character(
    iDealer$Vessel.Name
  )
)
## VTR numbers should be character strings to maintain leading zeros
iDealer$VTR=as.character(
  iDealer$Vtr.Serial.No
  )
## Species names can be converted directly frim ITIS numbers
iDealer$SPECIES=NA
for(i in 1:nrow(iDealer)){
  itis=iDealer$Species.Itis[i]
  iDealer$SPECIES[i]=as.character(
    unique(
      species[
        which(
          species$ITIS==itis
        ),
        "AFS"
      ]
    )
  )
}
## Live weights should be reported as numbers
iDealer$WEIGHT=as.numeric(
  as.character(
    iDealer$Live.Weight
  )
)
## Dates should be converted to POSIX values
iDealer$DATE=ymd(
  as.character(
    iDealer$Date.Sold
    )
  )
## Convert permit numbers in the iVTR file to vessel names using the
## reference values available in the iDealer file
iVTR$VESSEL=NA
for(i in 1:nrow(iVTR)){
  if(iVTR$PERMIT[i]%in%iDealer$PERMIT){
    iVTR$VESSEL[i]=unique(
      iDealer$VESSEL[which(
        iDealer$PERMIT==iVTR$PERMIT[i]
      )]
    )
  } else {
    iVTR$VESSEL[i]=NA
  }
}
iVTR=subset(
  iVTR,
  is.na(iVTR$VESSEL)==FALSE
)
## Read in the audit selection file provided by TEEM FISH to double check the
## results
AUDIT=readWorksheetFromFile(
  file="../ClosedAreaComparisons/FY19/RawData/Copy of 20200604_EFP_EM2_selected_trips.xlsx",
  sheet=1
)
## Create a list of all potential VTR numbers in each data set
AV=c(
  unique(EM$VTR),
  unique(iDealer$VTR),
  unique(iVTR$VTR),
  unique(subset(
    AUDIT,
    AUDIT$AUDIT_SELECTED==1
  )$TRIP_ID)
)
## Remove duplicate values
AV=subset(
  AV,
  duplicated(AV)==FALSE
)
## Create an empty data frame to store comparisons
QC=data.frame(
  VTR_SERIAL=as.character(),
  AUDIT=as.logical(),
  VTR=as.logical(),
  EM=as.logical()
)
for(i in AV){
  qc=data.frame(
    VTR_SERIAL=as.character(i),
    AUDIT=i%in%AUDIT$TRIP_ID,
    VTR=i%in%iVTR$VTR,
    EM=i%in%EM$VTR
  )
  QC=rbind(QC,qc)
}
## After conversations with TEEM FISH staff, it was determined that one 
## group of trips were included erroneously (should be in FY18, not FY19)
## so those data should be removed from the EM dataframe
kill=as.character(
  subset(
    QC,
    QC$AUDIT==FALSE&QC$VTR==FALSE&QC$EM==TRUE
  )$VTR_SERIAL
)
EM=subset(
  EM,
  EM$VTR%in%kill==FALSE
)
## ---------------------------
## All data are standardized and ready for analysis
## ---------------------------
## Create a new data frame to store information about whether trips took place 
## inside closed areas or not
CA=data.frame(
  VTR=as.character(),
  LAT=as.numeric(),
  LON=as.numeric(),
  CAII=as.logical(),
  CL=as.logical(),
  WGOM=as.logical(),
  OUT=as.logical()
)
## Read in spatial data for all EM reviews
for(i in 1:nrow(EM)){
  vtr=EM$VTR[i]
  lat=EM$STARTLAT[i]
  lon=EM$STARTLON[i]
  new=data.frame(
    VTR=as.character(vtr),
    LAT=as.numeric(lat),
    LON=as.numeric(lon),
    CAII=NA,
    CL=NA,
    WGOM=NA,
    OUT=NA
  )
  CA=rbind(CA,new)
  rm(new)
}
## Read in spatial data for all VTRs
for(i in 1:nrow(iVTR)){
  vtr=iVTR$VTR[i]
  lat=iVTR$LAT[i]
  lon=iVTR$LON[i]
  new=data.frame(
    VTR=as.character(vtr),
    LAT=as.numeric(lat),
    LON=as.numeric(lon),
    CAII=NA,
    CL=NA,
    WGOM=NA,
    OUT=NA
  )
  CA=rbind(CA,new)
  rm(new)
}
## Remove all duplicate entries from the spatial data frame
CA$dup=duplicated(CA)
CA=subset(
  CA,
  CA$dup==FALSE
)
CA$dup=NULL
## Remove all trips with a malfunctioning GPS
CA=subset(
  CA,
  CA$LAT>20 & abs(CA$LON)>20
)
## Add a column to record whether a trip was audited or not
CA$EM=NA
## Add a column to record the cod stock area impacted by the trip
CA$STOCK=NA
## Download groundfish closures from the NOAA website as a .zip archive of 
## shapefiles into a temporary file
dest_file="AllCA.zip"
urlzip="https://s3.amazonaws.com/media.fisheries.noaa.gov/2020-09/Groundfish_Closure_Areas_20180409_0.zip?ON7sHgWHiJxpWm.B1IW5REVNRKhUvMrz"
download.file(
  url=urlzip,
  destfile=dest_file,
  mode="wb"
)
zip::unzip(
  zipfile=dest_file,
  exdir="AllCA"
)
## Read in the shapefile that contains all closed areas
AllCA=readOGR(
  dsn="AllCA/Groundfish_Closure_Areas/Groundfish_Closure_Areas.shp"
)
## Convert the shapefile from NAD83 (the default projection from NOAA) to
## WGS84 (the default projection for GEBCO, for plotting over bathymetry)
AllCA=spTransform(
  AllCA, 
  CRS("+init=epsg:4326")
  )
## Break up the shapefile into individual closed areas
## In FY2019, the closed area list includes the following:
## Cashes Ledge Closure
CL=AllCA[AllCA$AREANAME=="Cashes Ledge Closure Area",]
## Closed Area II
CA2=AllCA[AllCA$AREANAME=="Closed Area II Closure Area",]
## Western Gulf of Maine
WGOM=AllCA[AllCA$AREANAME=="Western Gulf of Maine Closure Area",]

## Remove the temporary files from the directory
unlink("AllCA.zip")
unlink(
  "AllCA",
  recursive=TRUE
  )
## Download groundfish stock areas from the NOAA website as a .zip archive  
## of shapefiles into a temporary file
dest_file="AllStocks.zip"
urlzip="https://www.fisheries.noaa.gov/webdam/download/97266468"
download.file(
  url=urlzip,
  destfile=dest_file,
  mode="wb"
)
zip::unzip(
  zipfile=dest_file,
  exdir="AllStocks"
)
## Read in the shapefile that contains all closed areas
AllStocks=readOGR(
  dsn="AllStocks/Stock_Areas/Stock_Areas.shp"
)
## Convert the shapefile from NAD83 (the default projection from NOAA) to
## WGS84 (the default projection for GEBCO, for plotting over bathymetry)
AllStocks=spTransform(
  AllStocks, 
  CRS("+init=epsg:4326")
)
## Break up the shapefile into individual stock areas and select out cod 
## stock areas
## In FY2019, the cod stock area list includes the following:
## Gulf of Maine Cod
GOM=AllStocks[AllStocks$AREANAME=="GOM Cod Stock Area",]
## Georges Bank Cod
GB=AllStocks[AllStocks$AREANAME=="GB Cod Stock Area",]
## Separate out other stock areas as well
## Cape Cod / Gulf of Maine Yellowtail Flounder
CCGOM_YT=AllStocks[AllStocks$AREANAME=="CC/GOM Yellowtail Flounder Stock Area (BDAS)"]
## Southern New England / Mid Atlantic Yellowtail Flounder
SNEMA_YT=AllStocks[AllStocks$AREANAME=="SNE/MA Yellowtail Flounder Stock Area (BDAS)"]
## Southern New England / Mid Atlantic Winter Flounder
SNEMA_WF=AllStocks[AllStocks$AREANAME=="SNE/MA Winter Flounder Stock Area"]
## Georges Bank Yellowtail Flounder
GB_YT=AllStocks[AllStocks$AREANAME=="GB Yellowtail Flounder Stock Area"]
## Georges Bank Winter Flounder
GB_WF=AllStocks[AllStocks$AREANAME=="GB Winter Flounder Stock Area"]
## NOTE: Pollock, White Hake, Witch Flounder, and American Plaice are all
## single stocks
## NOTE: Haddock are managed in the same stock areas as Cod (GOM and GB)
## Remove the temporary files from the directory
unlink("AllStocks.zip")
unlink(
  "AllStocks",
  recursive=TRUE
)
## Create an empty column for cod stock area
CA$STOCK=NA
for(i in 1:nrow(CA)){
  ## Separate the individual trip out
  trip=CA[i,]
  ## Assign the trip a spatial reference
  coordinates(trip)=~LON+LAT
  ## Reproject the trip to the same coordinate system as the closed area 
  ## shapefiles
  proj4string(trip)="+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"
  trip=spTransform(
    trip, 
    CRS("+init=epsg:4326")
  )
  ## Check to see if each trip overlaps with the boundaries of Cashes Ledge
  if(is.na(over(trip,CL)$AREANAME)==TRUE){
    CA$CL[i]=FALSE
  } else {
    CA$CL[i]=ifelse(
      as.character(over(trip,CL)$AREANAME)=="Cashes Ledge Closure Area",
      TRUE,
      FALSE
    )
  }
  ## Check to see if each trip overlaps with the boundaries of Closed Area II
  if(is.na(over(trip,CA2)$AREANAME)==TRUE){
    CA$CAII[i]=FALSE
  } else {
    CA$CAII[i]=ifelse(
      as.character(over(trip,CA2)$AREANAME)=="Closed Area II Closure Area",
      TRUE,
      FALSE
    )
  }
  ## Check to see if each trip overlaps with the boundaries of the WGOM
  if(is.na(over(trip,WGOM)$AREANAME)==TRUE){
    CA$WGOM[i]=FALSE
  } else {
    CA$WGOM[i]=ifelse(
      as.character(over(trip,WGOM)$AREANAME)=="Western Gulf of Maine Closure Area",
      TRUE,
      FALSE
    )
  }
  ## If the trips do not take place in a closed area, label them as OUT
  CA$OUT[i]=ifelse(
    CA$CAII[i]+CA$CL[i]+CA$WGOM[i]==0,
    TRUE,
    FALSE
  )
  ## Check to see if the trips are in the GB or GOM cod stock area
  if(is.na(over(trip,GOM)$AREANAME)==FALSE){
    CA$STOCK[i]=ifelse(
      as.character(over(trip,GOM)$AREANAME)=="GOM Cod Stock Area",
      "GOM",
      CA$STOCK[i]
    )
  }
  if(is.na(over(trip,GB)$AREANAME)==FALSE){
    CA$STOCK[i]=ifelse(
      as.character(over(trip,GB)$AREANAME)=="GB Cod Stock Area",
      "GB",
      CA$STOCK[i]
    )
  }
  ## Check each trip to see if it was audited by the EM program
  CA$EM=ifelse(
    CA$VTR%in%EM$VTR,
    TRUE,
    FALSE
  )
}
## Link each VTR in the CA table with a vessel name
CA$VESSEL=NA
for(i in 1:nrow(CA)){
  if(as.character(CA$VTR[i])%in%as.character(iVTR$VTR)){
    CA$VESSEL[i]=as.character(
      unique(
        iVTR$VESSEL[which(
          iVTR$VTR==as.character(
            CA$VTR[i]
          )
        )
        ]
      )
    )
  }
}
## If linking by VTR number fails, try linking by permit number
## Create a list of permit numbers from the EM data to augment the dealer data
empn=select(EM,VTR,VESSEL)
empn$PN=substr(
  x=EM$VTR,
  start=1,
  stop=6
)
empn$VTR=NULL
empn$dup=duplicated(empn)
empn=subset(
  empn,
  empn$dup==FALSE
  )
empn$dup=NULL
for(i in 1:nrow(CA)){
  if(
    is.na(CA$VESSEL[i])
  ){
    pn=substr(
      x=CA$VTR[i],
      start=1,
      stop=6
    )
    if(pn%in%iDealer$Vessel.Permit.No){
      CA$VESSEL[i]=unique(
        iDealer[which(
          iDealer$Vessel.Permit.No==pn
        ),]$VESSEL
      )
    } else {
      if(pn%in%empn$PN){
        CA$VESSEL[i]=empn$VESSEL[which(empn$PN==pn)]
      }
    }
  }
}
## Make a list of all unique vessels in the CA data frame
VESSEL=unique(CA$VESSEL)[order(
  unique(CA$VESSEL)
  )]
VESSEL=subset(VESSEL,is.na(VESSEL)==FALSE)
## Create a table to store trips by closed area and vessel
CAV=data.frame(
  VESSEL=as.character(),
  CAII=as.numeric(),
  CL=as.numeric(),
  WGOM=as.numeric(),
  OUT=as.numeric()
)
## For each vessel, total up how many trips it took inside and outside of the
## closed areas
for(i in 1:length(VESSEL)){
  v=VESSEL[i]
  x=subset(CA,CA$VESSEL==v)
  ca2=sum(x$CAII)
  cl=sum(x$CL)
  wg=sum(x$WGOM)
  o=sum(x$OUT)
  y=data.frame(
    VESSEL=v,
    CAII=ca2,
    CL=cl,
    WGOM=wg,
    OUT=o
  )
  CAV=rbind(CAV,y)
}
CAA=data.frame(
  VESSEL=as.character(),
  TRIPS=as.numeric(),
  AUDITED=as.numeric()
)
for(i in 1:length(VESSEL)){
  v=VESSEL[i]
  x=subset(CA,CA$VESSEL==v)
  trips=nrow(x)
  a=sum(x$EM)
  y=data.frame(
    VESSEL=v,
    TRIPS=trips,
    AUDITED=a
  )
  CAA=rbind(CAA,y)
}
## Cod Discard Records
## Create an empty data frame to store the records
CodDiscards=data.frame(
  VTR=as.character(),
  CD=as.numeric(),
  DISCARD_SOURCE=as.character()
)
## For each trip, pull EM data (if available). If not, use the VTR data.
for(i in 1:nrow(CA)){
  trip=CA$VTR[i]
  em=subset(EM,EM$VTR==trip)
  if(nrow(em)>0){
    coddisc=sum(
      subset(em,em$SPECIES=="ATLANTIC COD")$DiscardWeight,
      na.rm=TRUE
    )
    cd=data.frame(
      VTR=trip,
      CD=coddisc,
      DISCARD_SOURCE="EM"
    )
  } else {
    vtr=subset(iVTR,iVTR$VTR==trip)
    coddisc=sum(
      subset(vtr,vtr$SPECIES=="ATLANTIC COD")$DISCARDED,
      na.rm=TRUE
    )
    cd=data.frame(
      VTR=trip,
      CD=coddisc,
      DISCARD_SOURCE="VTR"
    )
  }
  CodDiscards=rbind(CodDiscards,cd)
}
## Remove duplicate records
CodDiscards$dup=duplicated(CodDiscards)
CodDiscards=subset(
  CodDiscards,
  CodDiscards$dup==FALSE
)
CodDiscards$dup=NULL
## Add in closed area info
CodDiscards$CAII=FALSE
CodDiscards$CL=FALSE
CodDiscards$WGOM=FALSE
CodDiscards$OUT=FALSE
for(i in 1:nrow(CodDiscards)){
  x=subset(CA,CA$VTR==CodDiscards$VTR[i])
  CodDiscards$CAII[i]=sum(x$CAII)>0
  CodDiscards$CL[i]=sum(x$CL)>0
  CodDiscards$WGOM[i]=sum(x$WGOM)>0
  CodDiscards$OUT[i]=sum(x$OUT)>0
}
CodDiscards$MIX=(CodDiscards$CAII+CodDiscards$CL+CodDiscards$WGOM+CodDiscards$OUT)>1
## Add in Vessel names and cod stock areas
CodDiscards$VESSEL=NA
CodDiscards$STOCK=NA
for(i in 1:nrow(CodDiscards)){
  CodDiscards$VESSEL[i]=subset(
    CA,
    CA$VTR==CodDiscards$VTR[i]
  )$VESSEL[1]
  CodDiscards$STOCK[i]=subset(
    CA,
    CA$VTR==CodDiscards$VTR[i]
  )$STOCK[1]
}
## Add in landings values
CodDiscards$KALL=0
CodDiscards$HADDOCK=0
CodDiscards$COD=0
CodDiscards$LANDINGS_SOURCE=NA
for(i in 1:nrow(CodDiscards)){
  l=subset(
    iDealer,
    iDealer$Vtr.Serial.No==CodDiscards$VTR[i]
  )
  if(nrow(l)==0){
    v=subset(
      iVTR,
      iVTR$VTR==CodDiscards$VTR[i]
    )
    if(nrow(v)==0){
      v=subset(
        EM,
        EM$VTR==CodDiscards$VTR[i]
      )
      start=floor_date(
        min(v$STARTTIME),
        unit="day"
      )
      end=ceiling_date(
        max(v$ENDTIME),
        unit="day"
      )
    } else {
      start=floor_date(
        min(v$SAILDATE),
        unit="day"
        )
      end=ceiling_date(
        max(v$LANDDATE),
        unit="day"
      )
    }
    ves=v$VESSEL[1]
    l=subset(
      iDealer,
      iDealer$VESSEL==ves&iDealer$DATE%in%seq(
        from=start,
        to=end+86400,
        by="day"
      )
    )
  }
  CodDiscards$KALL[i]=sum(
    l$WEIGHT,
    na.rm=TRUE
    )
  CodDiscards$HADDOCK[i]=sum(
    subset(
      l,
      l$SPECIES=="HADDOCK"
    )$WEIGHT,
    na.rm=TRUE
  )
  CodDiscards$COD[i]=sum(
    subset(
      l,
      l$SPECIES=="ATLANTIC COD"
    )$WEIGHT,
    na.rm=TRUE
  )
  if(nrow(l)>0){
    CodDiscards$LANDINGS_SOURCE[i]="DEALER"
  }
}
## For records that still have blanks because the dealer data can't be
## linked or isn't available, use the self-reported VTR data to compile
## landings
for(i in 1:nrow(CodDiscards)){
  if(CodDiscards$KALL[i]==0){
    x=subset(
      iVTR,
      iVTR$VTR==CodDiscards$VTR[i]
    )
    if(nrow(x)>0){
      ## Calculcate KALL
      CodDiscards$KALL[i]=sum(
        x$KEPT,
        na.rm=TRUE
      )
      ## Calculate HADDOCK landings
      CodDiscards$HADDOCK[i]=sum(
        subset(
          x,
          x$SPECIES=="HADDOCK"
        )$KEPT
      )
      ## Calculate COD landings
      CodDiscards$COD[i]=sum(
        subset(
          x,
          x$SPECIES=="ATLANTIC COD"
        )$KEPT
      )
      ## Label the record
      CodDiscards$LANDINGS_SOURCE[i]="VTR"
    }
  }
}
## Analyze only those trips that report cod landings or discards
CodD=subset(
  CodDiscards,
  (CodDiscards$CD+CodDiscards$COD)>0
)
## Flag trips without landings data
CodD$LANDINGS=ifelse(
  CodD$KALL==0,
  FALSE,
  TRUE
)
## Assign each trip to a closed area group for plotting purposes
CodD$CA=ifelse(
  CodD$CAII==TRUE,
  "CAII",
  ifelse(
    CodD$CL==TRUE,
    "CL",
    ifelse(
      CodD$WGOM==TRUE,
      "WGOM",
      "OUT"
    )
  )
)
## Assign gear type to each trip
CodD$GEAR=NA
for(i in 1:nrow(CodD)){
  g=unique(
    subset(
      iVTR,
      iVTR$VTR==CodD$VTR[i]
    )$GEAR
  )
  if(length(g)==1){
    CodD$GEAR[i]=unique(g)
  }
  if(length(g)>1){
    g=g[order(g)]
    g=paste0(g,collapse="/")
    CodD$GEAR[i]=g
  }
}
## Add Haddock:Cod ratio for Longliners
CodD$HADRAT=NA
for(i in 1:nrow(CodD)){
  if(is.na(CodD$GEAR[i])==FALSE){
    if(CodD$GEAR[i]=="LONGLINE"){
      CodD$HADRAT[i]=ifelse(
        (CodD$CD[i]+CodD$COD[i])==0,
        CodD$HADDOCK[i],
        CodD$HADDOCK[i]/(CodD$CD[i]+CodD$COD[i])
      )
    }
  }
}
```


This report covers trips taken between `r ceiling_date(min(iVTR$SAILDATE),unit="day")` and `r floor_date(max(iVTR$LANDDATE),unit="day")`. 

*Table 1a -- Trips taken exclusively in closed areas by vessels participating in the electronic monitoring program. Closed area abbreviations are as follows: Closed Area II (CAII), Cashes Ledge (CL), Western Gulf of Maine (WGOM). Trips in the "Out" column took place outside of closed areas.* 
  
```{r echo=FALSE, results='asis'}
CAV$
options(digits=0)
as.data.frame(CAV) %>% regulartable() %>% autofit()
```

*Table 1b -- Total trips taken and total trips audited through the EM program by each vessel in FY19. The target audit rate this year is 50%*
  
```{r echo=FALSE, results='asis'}
options(digits=0)
as.data.frame(CAA) %>% regulartable() %>% autofit()
```

```{r Mapping, include=FALSE}
## Create a vector of blues for plotting a map of trips
blues=c(
  "lightsteelblue4", 
  "lightsteelblue3", 
  "lightsteelblue2", 
  "lightsteelblue1"
  )
## Create a vector of grays for plotting a map of trips
grays=c(
  gray(0.6), 
  gray(0.93), 
  gray(0.99)
  )
## Download bathymetric data for plotting trip locations
basemap=getNOAA.bathy(
  lon1=-75, 
  lon2=-65, 
  lat1=40, 
  lat2=48, 
  resolution=1
  )
```
```{r Map Plotting, echo=FALSE, fig.height=3.5}
## Plot the bathymetric map as a background
plot(
  basemap, 
  image = TRUE, 
  land = TRUE, 
  deep=-100000, 
  shallow=0, 
  step=999999, 
  drawlabels = FALSE, 
  bpal = list(
    c(
      min(basemap,na.rm=TRUE),
      0, 
      blues
      ), 
    c(
      0, 
      max(basemap, na.rm=TRUE), 
      grays
      )
    ), 
  lwd = 0.1
  )
## Add a point for each of the trips (red for unaudited, blue for audited)
points(
  x=subset(CA,CA$EM==FALSE)$LON,
  y=subset(CA,CA$EM==FALSE)$LAT,
  pch=22,
  col='black',
  bg='red'
  )
points(
  x=subset(CA,CA$EM==TRUE)$LON,
  y=subset(CA,CA$EM==TRUE)$LAT,
  pch=22,
  col='black',
  bg='blue'
)
## Add the Closed Area shapes
plot(
  AllCA,
  lwd=1,
  add=TRUE,
  dens=25,
  angle=45,
  col='yellow'
)
plot(
  AllCA,
  lwd=3,
  add=TRUE,
  border='yellow'
)
```

*Figure 1 -- Approximate location of fishing efforts under the EFP. Blue trips were audited and red trips were not. Closed areas are represented by yellow hatched polygons. NOTE: Some trips are plotting on shore, and it is unclear why, as both fishermen reported trips and GPS-recorded trips show up on land.*  
  
**COD DISCARDS**    
  
Overall, participating vessels have discarded `r sum(CodD$CD)` pounds of cod including `r sum(subset(CodD,CodD$CL==TRUE)$CD)` pounds in the Cashes Ledge closure, `r sum(subset(CodD,CodD$WGOM==TRUE)$CD)` pounds in the Western Gulf of Maine closure, `r sum(subset(CodD,CodD$CAII==TRUE)$CD)` pounds in Closed Area II, and `r sum(CodD$CD)-sum(c(subset(CodD,CodD$CL==TRUE)$CD,subset(CodD,CodD$WGOM==TRUE)$CD,subset(CodD,CodD$CAII==TRUE)$CD))` pounds outside of the closed areas.

```{r Discard Plotting, echo=FALSE, fig.height=3.5}
options(digits=2)
barplot(
  c(
    sum(
      subset(CodD,CodD$CL==TRUE)$CD
      ),
    sum(
      subset(CodD,CodD$WGOM==TRUE)$CD
      ),
    sum(
      subset(CodD,CodD$CAII==TRUE)$CD
      ),
    sum(
      CodD$CD
      )-sum(
        c(subset(CodD,CodD$CL==TRUE)$CD,
          subset(CodD,CodD$WGOM==TRUE)$CD,
          subset(CodD,CodD$CAII==TRUE)$CD)
      )
    ),
  names=c(
    "CL",
    "WGOM",
    "CAII",
    "OUT"
  ),
  width=c(
    1,
    1,
    1,
    1
  ),
  ylab="Total Discards (lbs)",
  col="lightgray"
)
```

*Figure 2 -- Barplot of total cod discards (lbs) by closed area from electronic monitoring project participants between `r ceiling_date(min(iVTR$SAILDATE),unit="day")` and `r floor_date(max(iVTR$LANDDATE),unit="day")`. Closed area abbreviations are as follows: Cashes Ledge (CL), Western Gulf of Maine (WGOM), and Closed Area II (CAII). Trips in the "OUT" column took place outside of closed areas. For the purposes of this analysis, if a trip took place even partially within a closed area it was counted towards that closed area's total.* 

Average per trip cod discards were `r round(sum(subset(CodD,CodD$CL==TRUE)$CD)/nrow(subset(CodD,CodD$CL==TRUE)),1)` pounds in the Cashes Ledge closure, `r round(sum(subset(CodD,CodD$WGOM==TRUE)$CD)/nrow(subset(CodD,CodD$WGOM==TRUE)),1)` pounds in the Western Gulf of Maine closure, `r round(sum(subset(CodD,CodD$CAII==TRUE)$CD)/nrow(subset(CodD,CodD$CAII==TRUE)),1)` pounds in Closed Area II, and `r round(sum(subset(CodD,CodD$OUT==TRUE)$CD)/nrow(subset(CodD,CodD$OUT==TRUE)),0)` pounds outside of the closed areas.

```{r Discard Plotting 2, echo=FALSE, fig.height=3.5}
boxplot(
  (CodD$CD/CodD$KALL)~CodD$CA,
  names=c(
    "CAII",
    "CL",
    "WGOM",
    "OUT"
  ),
  ylab="Cod Discards / KAll",
  col="lightgray",
  ylim=c(
    0,
    0.15
  ),
  xlab=""
)
```

*Figure 3 -- Box plot of per trip cod discards (lbs) divided by total landings for all species (KAll, lbs) inside and outside of closed areas by electronic monitoring program participants during FY2019. Closed area abbreviations are as follows: Closed Area II (CAII), Cashes Ledge (CL), Western Gulf of Maine (WGOM). Trips in the "OUT" column took place entirely outside of closed areas. For the purposes of this analysis, if a trip took place even partially within a closed area it was counted towards that closed area's total. Additionally, all trips that did not report groundfish landings (e.g., monkfish trips) were discarded from this analysis. A discard mortality rate of 1.0 was assumed across all gear types.*

**COD MORTALITY**

Cod mortality includes discards and landings. Discard mortality was assumed at 100% for this analysis.

```{r Figure 4, echo=FALSE, fig.height=3.5}
boxplot(
  ((CodD$CD+CodD$COD)/CodD$KALL)~CodD$CA,
  names=c(
    "CAII",
    "CL",
    "WGOM",
    "OUT"
  ),
  ylab="Cod Mortality / KAll",
  col="lightgray",
  xlab=""
)
```

*Figure 4 -- Box plot of per trip cod mortality (lbs of discards and landings) divided by total landings for all species (KAll, lbs) inside and outside of closed areas by electronic monitoring program participants during FY2019. Closed area abbreviations are as follows: Closed Area II (CAII), Cashes Ledge (CL), Western Gulf of Maine (WGOM). Trips in the "OUT" column took place entirely outside of closed areas. For the purposes of this analysis, if a trip took place even partially within a closed area it was counted towards that closed area's total. Additionally, all trips that did not report groundfish landings (e.g., monkfish trips) were discarded from this analysis. A discard mortality rate of 1.0 was assumed across all gear types.*
  
**COD LANDINGS**

Overall, participating vessels have landed `r sum(CodD$COD)` pounds of cod including `r sum(subset(CodD,CodD$CA=="CL")$COD)` pounds in the Cashes Ledge closure, `r sum(subset(CodD,CodD$CA=="WGOM")$COD)` pounds in the Western Gulf of Maine closure, `r sum(subset(CodD,CodD$CA=="CAII")$COD)` pounds in Closed Area II, and `r sum(subset(CodD,CodD$CA=="OUT")$COD)` pounds outside of the closed areas.

```{r Landing Plotting, echo=FALSE, fig.height=3.5}
barplot(
  c(
    sum(subset(CodD,CodD$CL==TRUE)$COD),
    sum(subset(CodD,CodD$WGOM==TRUE)$COD),
    sum(subset(CodD,CodD$CAII==TRUE)$COD),
    sum(subset(CodD,CodD$OUT==TRUE)$COD)
  ),
  names=c(
    "CL",
    "WGOM",
    "CAII",
    "Out"
    ),
  width=c(
    1,
    1,
    1,
    1),
  ylab="Landed (Live lbs)",
  col="lightgray"
  )
```

*Figure 6 -- Barplot of total cod landings (lbs) by closed area from electronic monitoring project participants between`r ceiling_date(min(iVTR$SAILDATE),unit="day")` and `r floor_date(max(iVTR$LANDDATE),unit="day")`. Closed area abbreviations are as follows: Closed Area II (CAII), Cashes Ledge (CL), Western Gulf of Maine (WGOM). Trips in the "OUT" column took place entirely outside of closed areas. For the purposes of this analysis, if a trip took place even partially within a closed area it was counted towards that closed area's total.* 
  
**SPECIES SPECIFIC COMPARISONS**
  
_**Longlined Haddock**_

Longline vessels participating in this project target haddock, fishing in open areas and in the Western Gulf of Maine closure. Assuming a 100% discard mortality rate, inside the Western Gulf of Maine closure, harvesters averaged `r mean(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$WGOM==TRUE)$HADRAT)` ± `r sd(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$WGOM==TRUE)$HADRAT)` pounds of haddock landed per pound of cod killed (as discards or landings) (median: `r mean(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$WGOM==TRUE)$HADRAT)`). 

```{r HadRat Plotting1, echo=FALSE, fig.height=3.5}
z=subset(CodD,CodD$GEAR=="LONGLINE"&CodD$WGOM==TRUE)$HADRAT
z=ifelse(
  z>1000,
  1000,
  z
)
hist(
  z,
  xlab="Haddock Landed : Cod Mortality Ratio (DM=100%)",
  main='',
  ylab="Trips",
  xlim=c(
    0,
    1000
  ),
  col='darkgray',
  breaks=seq(0,1000,10)
  )
abline(
  v=mean(
    z
  ),
  lty=1
)
abline(
  v=median(
    z
  ),
  lty=2
)
```

*Figure 7a -- Histogram of haddock landed (lbs) : cod mortality (lbs of discards + lbs of landings) ratio for longline vessels fishing in the Western Gulf of Maine closed area. The average ratio (solid black line) was `r mean(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$WGOM==TRUE)$HADRAT)` lbs of haddock landed per pound of cod mortality and the median ratio (dashed black line) was `r median(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$WGOM==TRUE)$HADRAT)` lbs of haddock landed per pound of cod mortality. NOTE: All ratios > 1000 lbs of haddock per lb of cod are displayed as 1000 in this graph.*
  
Assuming a 100% discard mortality rate, in the Cashes Ledge closure area, harvesters averaged `r mean(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$CL==TRUE)$HADRAT)` ± `r sd(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$CL==TRUE)$HADRAT)` pounds of haddock landed per pound of cod killed(as discards or landings)  (median: `r median(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$CL==TRUE)$HADRAT)`). 

```{r HadRat Plotting2, echo=FALSE, fig.height=3.5}
z=subset(CodD,CodD$GEAR=="LONGLINE"&CodD$CL==TRUE)$HADRAT
z=ifelse(
  z>1000,
  1000,
  z
)
hist(
  z,
  xlab="Haddock Landed : Cod Mortality Ratio (DM=100%)",
  main='',
  ylab="Trips",
  xlim=c(
    0,
    1000
  ),
  col='darkgray',
  breaks=seq(0,1000,10)
  )
abline(
  v=mean(
    z
  ),
  lty=1
)
abline(
  v=median(
    z
  ),
  lty=2
)
```

*Figure 7b -- Histogram of haddock landed (lbs) : cod mortality (lbs of discards + lbs of landings) ratio for longline vessels fishing in the Cashes Ledge closed area. The average ratio (solid black line) was `r mean(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$CL==TRUE)$HADRAT)` lbs of haddock landed per pound of cod mortality and the median ratio (dashed black line) was `r median(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$CL==TRUE)$HADRAT)` lbs of haddock landed per pound of cod mortality. NOTE: All ratios > 1000 lbs of haddock per lb of cod are displayed as 1000 in this graph.*

Assuming a 100% discard mortality rate, in the open areas, harvesters averaged `r mean(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$OUT==TRUE)$HADRAT)` ± `r sd(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$OUT==TRUE)$HADRAT)` pounds of haddock landed per pound of cod killed(as discards or landings)  (median: `r median(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$OUT==TRUE)$HADRAT)`). 

```{r HadRat Plotting3, echo=FALSE, fig.height=3.5}
z=subset(CodD,CodD$GEAR=="LONGLINE"&CodD$OUT==TRUE)$HADRAT
z=ifelse(
  z>1000,
  1000,
  z
)
hist(
  z,
  xlab="Haddock Landed : Cod Mortality Ratio (DM=100%)",
  main='',
  ylab="Trips",
  xlim=c(
    0,
    1000
  ),
  col='darkgray',
  breaks=seq(0,1000,10)
  )
abline(
  v=mean(
    z
  ),
  lty=1
)
abline(
  v=median(
    z
  ),
  lty=2
)
```

*Figure 7c -- Histogram of haddock landed (lbs) : cod mortality (lbs of discards + lbs of landings) ratio for longline vessels fishing in open areas. The average ratio (solid black line) was `r mean(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$OUT==TRUE)$HADRAT)` lbs of haddock landed per pound of cod mortality and the median ratio (dashed black line) was `r median(subset(CodD,CodD$GEAR=="LONGLINE"&CodD$OUT==TRUE)$HADRAT)` lbs of haddock landed per pound of cod mortality. NOTE: All ratios > 1000 lbs of haddock per lb of cod are displayed as 1000 in this graph.*
  
  **2019 FISHING YEAR SUMMARY**
  
*Table 2 -- Cod landings and discards (lbs) by stock. Cod quota is managed in two stocks: Georges Bank (GB) and the Gulf of Maine (GOM)*

```{r CodStocks, echo=FALSE, results='asis'}
STOCK=c("GB","GOM")
DISCARDS=c(
  sum(
    subset(
      CodD,
      CodD$STOCK=="GB"
    )$CD
  ),
  sum(
    subset(
      CodD,
      CodD$STOCK=="GOM"
    )$CD
  )
)
LANDINGS=c(
  sum(
    subset(
      CodD,
      CodD$STOCK=="GB"
    )$COD
  ),
  sum(
    subset(
      CodD,
      CodD$STOCK=="GOM"
    )$COD
  )
)
Cod=data.frame(
  STOCK,
  DISCARDS,
  LANDINGS
)
options(digits=0)
Cod %>% regulartable() %>% autofit() %>% set_formatter_type(Cod,
fmt_double = "%.00f")
```